---
title: "STAT- S 670 EXPLORATORY DATA ANALYSIS FINAL PROJECT (CHICAGO CRIME ANALYSIS"
author: "VISWA SUHAAS PENUGONDA, HARSHA SAI GADE, MANASA GUDISE"
output: pdf_document
---

```{r}

```

## NEEDED LIBRARIES

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(broom)
library(grid)
library(mgcv)
library(data.table)
library(stringr)
library(reshape2)
library(MASS)
library(RCurl)
library(GGally)
```

# PRE- PROCESSING

## LOAD THE DATASET

```{r}
DATA= read.csv("/Users/suhaaspenugonda/Documents/IU/Spring Semester 2022-23/Exploratory Data Analysis/FINAL PROJECT/Crimes_2022.csv")
head(DATA)
```

```{r}
# UNDERSTANDING THE NAMES OF THE COLUMNS IN THE DATASET

names(DATA)

# REMOVING COLUMNS WHICH DO NOT HAVE SIGNIFICANCE IN UNDERSTANDING THE DATA IN THE DATASET

COL_DROP = c("Case.Number", "Description", "IUCR", "Beat", "Block", "Ward", "FBI.Code", "X.Coordinate", "Y.Coordinate", "Location")

DATA = DATA[,!(names(DATA) %in% COL_DROP)]

names(DATA)

summary(DATA)
```

## PLOT OF THE FREQUENCY DISTRIBUTION OF VARIOUS TYPES OF CRIMES IN CHICAGO

```{r fig.width=10}
ggplot(DATA, aes(y = Primary.Type)) + theme_minimal() + 
  geom_bar() + xlab("COUNT OF THE CRIMES") + ylab("TYPE OF THE CRIME") + ggtitle("FREQUENCY DISTRIBUTION OF VARIOUS TYPES OF CRIMES IN CHICAGO")

CRIMES = DATA$Primary.Type
CRIMES = as.data.frame(table(CRIMES))
CRIMES = CRIMES[order(CRIMES$Freq,decreasing = TRUE),]
CRIMES = CRIMES[CRIMES$Freq > 5000,]
```

## PLOT OF THE FREQUENCY DISTRIBUTION OF VARIOUS TYPES OF CRIMES BASED ON THE LOCATION IN CHICAGO

```{r fig.height=15, fig.width=15}
LOCATION_CRIME = DATA$Location.Description
LOCATION_CRIME = as.data.frame(table(LOCATION_CRIME))
LOCATION_CRIME = LOCATION_CRIME[order(LOCATION_CRIME$Freq,decreasing = TRUE),]
LOCATION_CRIME = LOCATION_CRIME[1:10,]

ggplot(LOCATION_CRIME,aes(x=LOCATION_CRIME,y=Freq)) +  geom_bar(stat="identity") + coord_flip()+ theme_minimal()+
  labs(y= 'FREQUENCY DISTRIBUTION',x ='LOCATIONS' , title = 'FREQUENCY DISTRIBUTION OF VARIOUS TYPES OF CRIMES BASED ON THE LOCATION IN CHICAGO')
```

## FILTER THE DATA AND PERFORM BIVARIATE ANALYSIS TO FIND THE RELATION BETWEEN THE TYPE OF CRIMES WITH RESPECT TO THE LOCATION

```{r fig.height=10, fig.width=15}
DATA_SUBSET1 = DATA[DATA$Location.Description %in% LOCATION_CRIME$LOCATION_CRIME,]
DATA_SUBSET1 = DATA_SUBSET1[DATA_SUBSET1$Primary.Type %in% CRIMES$CRIMES,]


DATA_SUBSET2 = filter(DATA[DATA$Location.Description %in% LOCATION_CRIME$LOCATION_CRIME,], Primary.Type == 'SEX OFFENSE' | Primary.Type == 'PROSTITUTION' | Primary.Type == 'OBSCENITY' | Primary.Type == 'CRIMINAL SEXUAL ASSUALT' | Primary.Type == 'CRIM SEXUAL ASSUALT')

DATA_SUBSET2$Primary.Type_2 = 'SEXUAL OFFENSE AND RELATED'

DATA_SUBSET2$Primary.Type = DATA_SUBSET2$Primary.Type_2 
DATA_SUBSET2 = subset(DATA_SUBSET2, select = -c(Primary.Type_2))

DATA_MERGED = bind_rows(DATA_SUBSET1 , DATA_SUBSET2)

ggplot(DATA_MERGED,aes(x = Primary.Type))+
  geom_bar()+
  facet_wrap(~ Location.Description,nrow = 2,ncol=5)+ 
  theme(axis.text.x = element_text(angle = 90)) + labs(x = "TYPE OF CRIME" , y = "FREQUENCY DISTRIBUTION OF THE CRIME") + ggtitle(" FACET BAR PLOT TO KNOW THE OCCURENCE OF VARIOUS CRIMES BASED ON THE LOCATION")
```

## SEPARATION THE DATE COLUMN TO MONTH, DATE, AND TIME

```{r}
DATA_MERGED$DATE = as.Date(DATA_MERGED$Date, format = "%m/%d/%Y")
DATA_MERGED$TIME = format(as.POSIXct(DATA_MERGED$DATE), format = "%H:%M:%S")
DATA_MERGED$MONTH_NUM = strftime(DATA_MERGED$DATE, "%m")

#CONVERT THE VALUE TO MONTH TO NUMERIC
DATA_MERGED$MONTH = as.numeric(format(DATA_MERGED$DATE, "%m"))
```

## ADD THE NAME OF THE MONTH FOR MORE SPECIFICITY

```{r}
DATA_MERGED$MONTH_NAME = month.name[DATA_MERGED$MONTH]
all(DATA_MERGED$MONTH_NAME %in% month.name)
DATA_MERGED$MONTH_NAME <- factor(DATA_MERGED$MONTH_NAME, levels=month.name)
```

## PLOT OF THE FREQUENCY DISTRIBUTION OF VARIOUS CRIMES BASED ON THE MONTH IN CHICAGO DURING 2022

```{r fig.width=10}
ggplot(DATA_MERGED, aes(x = MONTH_NAME)) + theme(axis.text.x = element_text(angle = 90))  + 
  geom_bar() + ylab("FREQUENCY DISTRIBUTION OF VARIOUS CRIMES") + ggtitle("FREQUENCY DISTRIBUTION OF VARIOUS CRIMES BASED ON THE MONTH IN CHICAGO DURING 2022") + xlab("NAME OF THE MONTH")
```

## ENCODE THE "Arrest", "Domestic" COLUMNS FOR FINAL OUTPUT

```{r}
# CONVERT 'Arrest' TO LOGICAL
DATA_MERGED$Arrest <- as.logical(DATA_MERGED$Arrest)

# CONVERT 'Domestic' TO LOGICAL
DATA_MERGED$Domestic <- as.logical(DATA_MERGED$Domestic)

# CONVERT 'Arrest' and 'Domestic' TO NUMERIC
DATA_MERGED$ARREST_NUM <- as.numeric(DATA_MERGED$Arrest)
DATA_MERGED$DOMESTIC_NUM <- as.numeric(DATA_MERGED$Domestic)

DATA_MERGED$ARREST_NUM[DATA_MERGED$Arrest == 'TRUE'] = 1
DATA_MERGED$ARREST_NUM[DATA_MERGED$Arrest == 'FALSE'] = 0

DATA_MERGED$DOMESTIC_NUM[DATA_MERGED$Domestic == 'TRUE'] = 1
DATA_MERGED$DOMESTIC_NUM[DATA_MERGED$Domestic == 'FALSE'] = 0
```

## ADD THE NAME OF THE DAY FOR MORE SPECIFICITY AND ENCODE THEM INTO NUMERIC VALUES

```{r}
DATA_MERGED$DAY <- weekdays(as.Date(DATA_MERGED$DATE))

DATA_MERGED = DATA_MERGED %>% mutate(DAY_NUM = recode(DAY, 'Monday' = 1, 'Tuesday' = 2,'Wednesday'=3,'Thursday'=4,'Friday'=5,'Saturday'=6,'Sunday'=7))
```

## PLOT OF THE FREQUENCY DISTRIBUTION OF VARIOUS CRIMES BASED ON THE DAY IN A WEEK IN CHICAGO DURING 2022

```{r fig.width=12}
ggplot(DATA_MERGED, aes(x = DAY)) + theme(axis.text.x = element_text(angle = 90))  + 
  geom_bar() + ylab("FREQUENCY DISTRIBUTION OF VARIOUS CRIMES") + ggtitle("PLOT OF THE FREQUENCY DISTRIBUTION OF VARIOUS CRIMES BASED ON THE DAY IN A WEEK IN CHICAGO DURING 2022") + xlab("NAME OF THE DAY")
```

## SAVE THE CHANGES MADE TO THE DATESET TO THE ACTUAL CSV FILE

```{r}
write.csv(DATA_MERGED,'PREPROCESSED_CHICAGO_CRIMES.csv')
```

## LOAD THE DATASET

```{r}
DATA= read.csv("/Users/suhaaspenugonda/Documents/IU/Spring Semester 2022-23/Exploratory Data Analysis/FINAL PROJECT/PREPROCESSED_CHICAGO_CRIMES.csv")

head(DATA)
```

## SORT THE DATA IN ASCENDING ORDER BASED ON THE OCCURANCE OF THE CRIME

```{r}
DATA <- DATA[order(DATA$Date), ]
DATA$MONTH_NAME = factor(DATA$MONTH_NAME , levels = month.name )
DATA$DAY = factor(DATA$DAY , levels= c("Sunday", "Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```

## PLOT OF THE FREQUENCY DISTRIBUTION OF THE CRIMES BASED ON THE LOCATION IN A MONTH

```{r fig.width=10}
ggplot(DATA,aes(x = MONTH_NAME))+
  geom_bar()+
  facet_wrap(~Location.Description)+
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("PLOT OF THE FREQUENCY DISTRIBUTION OF THE CRIMES BASED ON THE LOCATION IN A MONTH")
```

# ANALYSIS TO KNOW THE PROPORTION OF ARRESTS FOR DIFFERENT CRIMES AT VARIOUS LOCATIONS

## STACKED BAR PLOT BASED ON THE PROPORTION OF ARRESTS AT EACH CRIME LOCATION

```{r fig.height=10, fig.width=15}
ggplot(DATA,aes(x = Location.Description,fill = Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
  geom_bar(position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("ARREST PROPORTION") + xlab("LOCATION OF THE CRIME") + ggtitle("STACKED BAR PLOT BASED ON THE PROPORTION OF ARRESTS AT EACH CRIME LOCATION")
```

## STACKED BAR PLOT BASED ON THE PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES

```{r fig.height=7, fig.width=10}
ggplot(DATA,aes(x = Primary.Type,fill = Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES") + xlab("TYPE OF THE CRIME") + ggtitle("STACKED BAR PLOT BASED ON THE PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES")
```

## FACETED BAR PLOT BASED ON THE PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES AT VARIOUS LOCATIONS

```{r fig.height=10, fig.width=12}
ggplot(DATA,aes(x = Location.Description,fill = Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES") + xlab("LOCATION OF THE CRIME") + ggtitle("FACETED BAR PLOT BASED ON THE PROPORTION OF ARRESTS MADE FOR VARIOUS CRIMES AT VARIOUS LOCATIONS") + facet_wrap(~Primary.Type)
```

# ANALYSIS OF ARRESTS FOR DIFFERENT CRIMES ACROSS THE VARIOUS MONTHS/DAYS OF THE YEAR

## ANALYSIS FOR SEXUAL OFFENSE AND RELATED

```{r}
SEXUAL_RELATED = subset(DATA,Primary.Type=='SEXUAL OFFENSE AND RELATED')
```

### PLOT OF THE PROPORTION OF ARRESTS FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS MONTHS OF THE YEAR

```{r fig.width=12}
SEXUAL_RELATED_DATA1 <- SEXUAL_RELATED %>%
  group_by(MONTH_NAME,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(SEXUAL_RELATED_DATA1,aes(x = SEXUAL_RELATED_DATA1$MONTH_NAME,y= pct,fill = SEXUAL_RELATED_DATA1$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS MONTHS OF THE YEAR") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("MONTHS OF THE ARRESTS") + ggtitle("PROPORTION OF ARRESTS FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS MONTHS OF THE YEAR")
```

### PLOT OF THE PROPORTION OF ARREST FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK

```{r fig.width=12}
SEXUAL_RELATED_DATA2 <- SEXUAL_RELATED %>%
  group_by(DAY,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(SEXUAL_RELATED_DATA2,aes(x = SEXUAL_RELATED_DATA2$DAY,y= pct,fill = SEXUAL_RELATED_DATA2$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARREST FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("DAY OF THE CRIME OCCURANCE") + ggtitle("PROPORTION OF ARREST FOR SEXUAL OFFENSE AND RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK")
```

## ANALYSIS OF WEAPONS VIOLATION

```{r}
WEAPON_RELATED = subset(DATA,Primary.Type=='WEAPONS VIOLATION')
```

### PLOT OF THE PROPORTION OF ARRESTS FOR WEAPONS VIOLATION CRIMES ACROSS VARIOUS MONTHS OF THE YEAR

```{r fig.width=12}
WEAPON_RELATED_DATA1 <- WEAPON_RELATED %>%
  group_by(MONTH_NAME,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(WEAPON_RELATED_DATA1,aes(x = WEAPON_RELATED_DATA1$MONTH_NAME,y= pct,fill = WEAPON_RELATED_DATA1$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR WEAPON_RELATED CRIMES ACROSS VARIOUS MONTHS OF THE YEAR") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("MONTHS OF THE ARRESTS") + ggtitle("PROPORTION OF ARRESTS FOR WEAPON_RELATED CRIMES ACROSS VARIOUS MONTHS OF THE YEAR")
```

### PLOT OF THE PROPORTION OF ARREST FOR WEAPON_RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK

```{r fig.width=12}
WEAPON_RELATED_DATA2 <- WEAPON_RELATED %>%
  group_by(DAY,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(WEAPON_RELATED_DATA2,aes(x = WEAPON_RELATED_DATA2$DAY,y= pct,fill = WEAPON_RELATED_DATA2$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARREST FOR WEAPON_RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("DAY OF THE CRIME OCCURANCE") + ggtitle("PROPORTION OF ARREST FOR WEAPON_RELATED CRIMES ACROSS VARIOUS DAYS OF THE WEEK")
```

# ANALYSIS OF ARRESTS FOR DIFFERENT CRIMES ACROSS THE VARIOUS LOCATIONS

## ANALYSIS FOR STREETS

```{r}
STREETS = subset(DATA,Location.Description == "STREET")
```

### PLOT OF THE PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS MONTHS OF THE YEAR

```{r fig.width=12}
STREETS_DATA1 <- STREETS %>%
  group_by(MONTH_NAME,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(STREETS_DATA1,aes(x = STREETS_DATA1$MONTH_NAME,y= pct,fill = STREETS_DATA1$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS MONTHS OF THE YEAR") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("MONTHS OF THE ARRESTS") + ggtitle("PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS MONTHS OF THE YEAR")
```

### PLOT OF THE PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS DAYS OF THE WEEK

```{r fig.width=12}
STREETS_DATA2 <- STREETS %>%
  group_by(DAY,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(STREETS_DATA2,aes(x = STREETS_DATA2$DAY,y= pct,fill = STREETS_DATA2$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS DAYS OF THE WEEK") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("DAY OF THE CRIME OCCURANCE") + ggtitle("PROPORTION OF ARRESTS FOR STREET CRIMES ACROSS VARIOUS DAYS OF THE WEEK")
```

## ANALYSIS FOR APARTMENTS

```{r}
APARTMENTS = subset(DATA,Location.Description == "APARTMENT")
```

### PLOT OF THE PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS MONTHS OF THE YEAR

```{r fig.width=12}
APARTMENTS_DATA1 <- APARTMENTS %>%
  group_by(MONTH_NAME,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(APARTMENTS_DATA1,aes(x = APARTMENTS_DATA1$MONTH_NAME,y= pct,fill = APARTMENTS_DATA1$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS MONTHS OF THE YEAR") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("MONTHS OF THE ARRESTS") + ggtitle("PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS MONTHS OF THE YEAR")
```

### PLOT OF THE PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS DAYS OF THE WEEK

```{r fig.width=12}
APARTMENTS_DATA2 <- APARTMENTS %>%
  group_by(DAY,Arrest) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

ggplot(APARTMENTS_DATA2,aes(x = APARTMENTS_DATA2$DAY,y= pct,fill = APARTMENTS_DATA2$Arrest)) + theme(axis.text.x = element_text(angle = 90)) +
    geom_bar(stat = "identity",
           position = "fill") +scale_fill_manual(values = c("grey", "black"), 
                    name = "CRIME THAT LEAD TO AN ARREST") + ylab("PROPORTION OF ARRESTS") + xlab("LOCATION OF THE CRIME") + ggtitle("PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS DAYS OF THE WEEK") + geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) + ylab("PROPORTION OF ARRESTS") + xlab("DAY OF THE CRIME OCCURANCE") + ggtitle("PROPORTION OF ARRESTS FOR APARTMENT CRIMES ACROSS VARIOUS DAYS OF THE WEEK")
```

# MODELLING TO GET A BETTER ACCURACY OF FUTURE CRIME PREDICTION

```{r}
LOG_REG <- glm(ARREST_NUM ~ Primary.Type  + Location.Description , family= 'binomial',data = DATA)

summary(LOG_REG)
```

## CREATE A DATAFRAME TO STORE ALL POSSIBLE COMBINATIONS OF 'Primary.Type', 'Location.Description' & 'Domestic'

```{r}
CRIMES_DATA = expand.grid(Primary.Type = c("BATTERY", "BURGLARY", "CRIMINAL DAMAGE" , "DECEPTIVE PRACTICE" , "MOTOR VEHICLE THEFT", "OTHER OFFENSE" , "ROBBERY" , "SEXUAL OFFENSE AND RELATED", "THEFT" , "WEAPONS VIOLATION"), Location.Description = c("APARTMENT","COMMERCIAL / BUSINESS OFFICE", "OTHER (SPECIFY)", "PARKING LOT / GARAGE (NON RESIDENTIAL)","RESIDENCE","RESTAURANT", "SIDEWALK" , "SMALL RETAIL STORE" , "STREET") , Domestic  = c("TRUE","FALSE"))
```

## PREDICT THE ARREST PROBABILITIES AGAINST THE NEW DATA AND THE ORIGINAL DATA

```{r}
# Predicting the arrest probabilities of 'CRIMES_DATA' using the logistic regression model and storing the predicted probabilities in a new column ARREST_PROB in CRIMES_DATA.

CRIMES_DATA$ARREST_PROB = predict(LOG_REG, CRIMES_DATA, type="response")

# Predicting the arrest probabilities of the original crimes data 'DATA' using the same logistic regression model and storing the predicted probabilities in a new column 'PROBABILITIES' in DATA.

DATA$PROBABILITIES = predict(LOG_REG, DATA, type="response")
```

```{r}

# The ifelse() function is used to create a new column called ARREST_NUM in the CRIMES_DATA data frame based on the ARREST_PROB column. The ARREST_NUM column will have a value of 1 if the corresponding ARREST_NUM value is greater than 0.5, indicating that an arrest is likely, and a value of 0 otherwise.

CRIMES_DATA$ARREST_NUM = ifelse(CRIMES_DATA$ARREST_PROB > 0.5,1,0)
```

## FACET PLOT OF THE ACTUAL VS THE PROBABLE ARRESTS AT VARIOUS LOCATIONS

```{r fig.height=6, fig.width=10}
ggplot(DATA,aes(y =Primary.Type))+geom_point(aes(x=PROBABILITIES,color=Arrest), size=3 , alpha = 0.7) + scale_color_manual(values = c("TRUE" = "black", "FALSE" = "brown"))+ facet_wrap(~DATA$Location.Description) + xlab("ARREST PROBABILITY") + ylab("TYPE OF THE CRIME")+ggtitle("FACET PLOT OF THE ACTUAL VS THE PROBABLE ARRESTS AT VARIOUS LOCATIONS") 
```

## FACET PLOT OF THE ACTUAL VS THE PROBABLE ARRESTS FOR VARIOUS CRIMES

```{r fig.height=6, fig.width=11}
ggplot(DATA,aes(y =Location.Description))+geom_point(aes(x=PROBABILITIES,color=Arrest), size=3 , alpha = 0.7) + scale_color_manual(values = c("TRUE" = "black", "FALSE" = "brown"))+ facet_wrap(~DATA$Primary.Type) + xlab("ARREST PROBABILITY") + ylab("LOCATION OF THE CRIME")+ggtitle("FACET PLOT OF THE ACTUAL VS THE PROBABLE ARRESTS FOR VARIOUS CRIMES") 
```








